---
title: "goalie"
date: "`r BiocStyle::doc_date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{goalie}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
abstract: Assertive check functions for defensive R programming.
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

# Introduction

[goalie][] is an attempt to incorporate elements of multiple assertive check packages into a single package with as few dependencies as possible. All assertive checks are written in the most basic [R][] code possible, without reliance on compilation (e.g. C++/Rcpp). It is still a work in progress, and feature requests are welcome.

R exports the `stopifnot()` function as a simple assert engine that supports multiple boolean returns (`TRUE`/`FALSE`) in a single call, separated by commas. This is a nice, flexible system for defensive programming. However, `stopifnot()` does not currently support custom error messages, and can be confusing to debug.

Similar to the approach used in the [assertthat][] package, goalie exports the `assert()` function, which is designed to be a drop-in replacement for `stopifnot()` with improved error message support. `validate()` is a variation on `assert()` that is intended to be used within S4 class validity checks, which is useful for [Bioconductor][] packages.

Additionally, goalie exports a number of assertive check functions that are intended to harden bioinformatics-oriented functions in R. These check functions return `logical` with an additional `"goalie"` S3 class definition, which enables us to define custom print methods that display additional useful information about the cause of the error (see `cause()` and `print()` for details). This approach draws heavily from the [assertive][] package.

The goalie source code incorporates elements from these excellent assert check packages:

- [assertive][] by Richie Cotton.
- [assertthat][] by Hadley Wickham.
- [checkmate][] by Michael Lang.

# The assert check engine

Todo

# Checking function inputs

Todo

# Creating new assert checks

Todo

# S4 class validity checks

Todo

# R session information

```{r}
utils::sessionInfo()
```

[assertive]: https://cran.r-project.org/package=assertive
[assertthat]: https://cran.r-project.org/package=assertthat
[BiocManager]: https://cran.r-project.org/package=BiocManager
[bioconda]: https://bioconda.github.io/
[Bioconductor]: https://bioconductor.org/
[checkmate]: https://cran.r-project.org/package=checkmate
[conda]: https://conda.io/
[goalie]: https://goalie.acidgenomics.com/
[R]: https://www.r-project.org/
